% ============================================================
%  Bibliographie
% ============================================================
\chapter*{Références bibliographiques}
\addcontentsline{toc}{chapter}{Références bibliographiques}
\markboth{Références bibliographiques}{Références bibliographiques}

\begin{thebibliography}{99}

% --- Logique des défauts ---
\bibitem{reiter1980}
  R.~Reiter,
  \textit{A Logic for Default Reasoning},
  Artificial Intelligence, vol.~13, n\textsuperscript{o}~1--2,
  pp.~81--132, 1980.

% --- Logique de description ---
\bibitem{baader2003}
  F.~Baader, D.~Calvanese, D.~McGuinness, D.~Nardi, P.~Patel-Schneider (éds.),
  \textit{The Description Logic Handbook: Theory, Implementation and Applications},
  Cambridge University Press, 2003.

% --- HermiT ---
\bibitem{hermit}
  R.~Shearer, B.~Motik, I.~Horrocks,
  \textit{HermiT: A Highly-Efficient OWL Reasoner},
  Proc. of the 5th International Workshop on OWL: Experiences and Directions (OWLED),
  Karlsruhe, Germany, 2008.

% --- OWL 2 ---
\bibitem{owl2}
  W3C OWL Working Group,
  \textit{OWL 2 Web Ontology Language — Document Overview},
  W3C Recommendation, 27 October 2009.
  \url{https://www.w3.org/TR/owl2-overview/}

% --- OWL API ---
\bibitem{owlapi}
  M.~Horridge, S.~Bechhofer,
  \textit{The OWL API: A Java API for OWL Ontologies},
  Semantic Web, vol.~2, n\textsuperscript{o}~1, pp.~11--21, 2011.

% --- TweetyProject ---
\bibitem{tweety}
  M.~Thimm,
  \textit{Tweety — A Comprehensive Collection of Java Libraries for
  Logical Aspects of Artificial Intelligence and Knowledge Representation},
  Proc. of the 14th International Conference on Principles of Knowledge
  Representation and Reasoning (KR), Vienna, Austria, 2014.
  \url{https://tweetyproject.org}

% --- Logique modale ---
\bibitem{kripke1963}
  S.~A.~Kripke,
  \textit{Semantical Analysis of Modal Logic I: Normal Modal Propositional Calculi},
  Zeitschrift für Mathematische Logik und Grundlagen der Mathematik,
  vol.~9, pp.~67--96, 1963.

\bibitem{blackburn2001}
  P.~Blackburn, M.~de Rijke, Y.~Venema,
  \textit{Modal Logic},
  Cambridge Tracts in Theoretical Computer Science, vol.~53,
  Cambridge University Press, 2001.

% --- Réseaux sémantiques ---
\bibitem{quillian1968}
  M.~R.~Quillian,
  \textit{Semantic Memory},
  in M.~Minsky (éd.), Semantic Information Processing,
  MIT Press, pp.~216--270, 1968.

% --- Logique classique ---
\bibitem{sat4j}
  D.~Le Berre, A.~Parrain,
  \textit{The SAT4J Library, Release 2.2},
  Journal on Satisfiability, Boolean Modeling and Computation (JSAT),
  vol.~7, pp.~59--64, 2010.
  \url{http://www.sat4j.org}

% --- RCR général ---
\bibitem{brachman2004}
  R.~J.~Brachman, H.~J.~Levesque,
  \textit{Knowledge Representation and Reasoning},
  Morgan Kaufmann Publishers, 2004.

% --- JavaFX ---
\bibitem{javafx}
  The OpenJFX Team,
  \textit{OpenJFX — JavaFX 17 Documentation},
  2021.
  \url{https://openjfx.io}

\end{thebibliography}
